---
title: "River Radii Paper: Results display"
author: "Francois Thoral"
format: html
self-contained: true
code-fold: true
number-sections: true
toc: true
toc-depth: 5
editor: visual
bibliography: references.bib
---

## Introduction and Methods

This document shows the R code and data used to generate the figures for the paper @orchard2025river.

This interactive document summarizes the relationships between the size of a buffer radius around river mouths and metrics of coastal water clarity KdPAR and Kd490 (mean, median, variance, 50-10-90-95th percentiles, % of valid observations), based on the 2013-2022 period particularly looking at the effect of the river mouth classes. We suggest a new grouping of classes too.

We run Kruskall-Wallis on pair-wise comparisons (Wilcoxon rank sum tests with continuity correction) to look for significant differences in distribution of Kd490 and KdPAR values between classes and for each buffer sizes.

**Data used:**

-   River names and river mouths location: `River Environmental Classification (REC2)` + Shane Orchard polish, January 2025 version.

-   Coastal water clarity:

    -   Downwelling light attenuation of PAR (400-700nm) coefficient (/m) [@gall2022satellite], version 5, July 2002 to August 2023. [`https://gis.niwa.co.nz/portal/apps/experiencebuilder/template/?id=9794f29cd417493894df99d422c30ec2`](https://gis.niwa.co.nz/portal/apps/experiencebuilder/template/?id=9794f29cd417493894df99d422c30ec2).

    -   Downwelling light attenuation at 490nm coefficient (/m) OC-CCI product, 4km resolution, <https://www.oceancolour.org/thredds/ncss/grid/CCI_ALL-v6.0-DAILY/dataset.html>, using QAA to calculate the IOPs, then Lee (2005) to get Kd490 from IOPs.

+------------------------+--------------------+
| **Classes**            |                    |
+========================+====================+
| Waituna-type lagoon    | 1.  Open Coast     |
+------------------------+--------------------+
| Hāpua-type lagoon      | 1.  Open Coast     |
+------------------------+--------------------+
| Beach streams          | 1.  Open Coast     |
+------------------------+--------------------+
| Freshwater river mouth | 1.  Open Coast     |
+------------------------+--------------------+
| Tidal river mouth      | 1.  Open Coast     |
+------------------------+--------------------+
| Tidal lagoon           | 1.  Open Coast     |
+------------------------+--------------------+
| Shallow drowned valley | 2.  Enclosed Coast |
+------------------------+--------------------+
| Deep drowned valley    | 2.  Enclosed Coast |
+------------------------+--------------------+
| Fjord                  | 2.  Enclosed Coast |
+------------------------+--------------------+

: Table 1: River mouth hydrosystem original classes (left) and suggested new grouping (right)

## Results

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(sf)
library(plotly)
library(viridis)
library(knitr)
library(broom)
library(DT)
library(ggpubr)
library(trend)
```

### Figure 5 Kd490, KdPAR boxplot, all radii

```{r}
#| message: false
#| warning: false
kd_metrics <- read_csv('KdPAR_SCENZ_v05_Kd490_OCCCI_10yrMetrics_From_1D_AllBuffers_Rivermouths_REC2_class2.csv') %>% 
      select(-CLASS,Interface)


rivermouth <- read_csv('REC2rivermouths_Order5-8_Reg_BioR_EcoD_HSclass_named.csv') %>% 
  mutate(NAME_name2 = paste0(NAME,'_',name_2)) %>% 
  select(NAME_name2,CLASS,BioR__BioR,CATCHMT_AR,REGC2017_N,StreamOrde,RID) %>% 
  mutate(CLASS2 = case_when(
    CLASS == "2. Waituna-type lagoon" | CLASS == "3. Hāpua-type lagoon" | CLASS == "4. Beach stream" | CLASS == "5. Freshwater river mouth" | CLASS == "6. Tidal river mouth" | CLASS == "7. Tidal lagoon" ~ 'Open Coast',   
    CLASS == "8. Shallow drowned valley" | CLASS == "9. Deep drowned valley" | CLASS == "10. Fjord" ~ 'Enclosed Coast',
    TRUE ~ 'NA')) %>% 
    mutate(CLASS = case_when(
    CLASS == "2. Waituna-type lagoon" ~ "Waituna-type lagoon",
    CLASS == "3. Hāpua-type lagoon" ~ "Hāpua-type lagoon",
    CLASS == "4. Beach stream" ~ "Beach stream",
    CLASS == "5. Freshwater river mouth" ~ "Freshwater river mouth",
    CLASS == "6. Tidal river mouth" ~ "Tidal river mouth",
    CLASS == "7. Tidal lagoon" ~ 'Tidal lagoon',   
    CLASS == "8. Shallow drowned valley" ~ "Shallow drowned valley",
    CLASS == "9. Deep drowned valley" ~ "Deep drowned valley",
    CLASS == "10. Fjord" ~ 'Fjord',
    TRUE ~ 'NA')) %>% 
    mutate(CLASS = factor(CLASS, levels = c("Waituna-type lagoon", 
                                          "Hāpua-type lagoon", 
                                          "Beach stream",
                                          "Freshwater river mouth", 
                                          "Tidal river mouth", 
                                          "Tidal lagoon",
                                          "Shallow drowned valley",
                                          "Deep drowned valley",
                                          "Fjord")))

kd_metrics <- inner_join(kd_metrics,rivermouth)

kd_variance <- kd_metrics %>% group_by(Buffer_Size) %>% 
  filter(Metrics == 'mean') %>% 
  mutate(variance = var(KdPAR,na.rm=T),
         Buffer_Size = Buffer_Size/1000) %>% 
  filter(!is.na(KdPAR))

ratio <- 15

p <- ggplot(data=kd_variance) +
  geom_boxplot(aes(x=factor(Buffer_Size),y=KdPAR)) +
  geom_path(aes(x= factor(Buffer_Size), y=variance*ratio), colour='red',group=1) +
  geom_point(aes(x= factor(Buffer_Size), y=variance*ratio), colour='red',group=1) +
  scale_y_continuous(
    name = "KdPAR (/m)",
    # Add a second axis and specify its features
    sec.axis = sec_axis(~ ./ratio, name="Variance (/m)^2")) +
  scale_fill_viridis_d() +
  xlab('Radius (km)') +
  theme_light() +
  theme(legend.position = 'bottom',
        axis.text.y.right = element_text(color = "red"),
        axis.line.y.right = element_line(color = "red"),
        axis.title.y.right = element_text(color = "red"))

kd_variance2 <- kd_metrics %>% group_by(Buffer_Size) %>% 
  filter(Metrics == 'mean') %>% 
  mutate(variance = var(Kd490,na.rm=T),
         Buffer_Size = Buffer_Size/1000) %>% 
  filter(!is.na(Kd490))

ratio <- 60

p2 <- ggplot(data=kd_variance2) +
  geom_boxplot(aes(x=factor(Buffer_Size),y=Kd490)) +
  geom_path(aes(x= factor(Buffer_Size), y=variance*ratio), colour='red',group=1) +
  geom_point(aes(x= factor(Buffer_Size), y=variance*ratio), colour='red',group=1) +
  scale_y_continuous(
    name = "Kd490 (/m)",
    # Add a second axis and specify its features
    sec.axis = sec_axis(~ ./ratio, name="Variance (/m)^2")) +
  scale_fill_viridis_d() +
  xlab('Radius (km)') +
  theme_light() +
  theme(legend.position = 'bottom',
        axis.text.y.right = element_text(color = "red"),
        axis.line.y.right = element_line(color = "red"),
        axis.title.y.right = element_text(color = "red"))

p_obs <- kd_metrics %>% 
  filter(Metrics == 'valid_days_pc') %>% 
  filter(!is.na(KdPAR)) %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  ggplot(data=.,aes(x=factor(Buffer_Size),y=KdPAR)) +
  geom_boxplot() +
  scale_fill_viridis_d() +
  #ggtitle('Percentage of valid observations of KdPAR (2013-2022)') +
  ylab('Percent of valid observations (%)') + xlab('Radius (km)') +
  theme_light() +
  theme(legend.position = 'bottom')

p2_obs <- kd_metrics %>% 
  filter(Metrics == 'valid_days_pc') %>% 
  filter(!is.na(Kd490)) %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  ggplot(data=.,aes(x=factor(Buffer_Size),y=Kd490)) +
  geom_boxplot() +
  scale_fill_viridis_d() +
  #ggtitle('Percentage of valid observations of Kd490 (2013-2022)') +
  ylab('Percent of valid observations (%)') + xlab('Radius (km)') +
  theme_light() +
  theme(legend.position = 'bottom')

figure5 <- ggarrange(p2,
                     p,
                     p2_obs,
                     p_obs,
                     labels = c("A", "B","C","D"),
                     ncol = 2, nrow = 2, align = 'v')
figure5
```

### KdPAR and Kd490 pairwise comparison

#### KdPAR Pairwise comparison

```{r}
aa <-  kd_metrics %>% 
  filter(!is.na(KdPAR)) %>% 
  filter(Metrics == 'mean') %>% 
  select(KdPAR,Buffer_Size) %>% 
  mutate(Buffer_Size = Buffer_Size/1000)


pw <- pairwise.wilcox.test(aa$KdPAR,aa$Buffer_Size)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('1',
                    '2',
                    '3',
                    '4',
                    '5',
                    '10',
                    '15'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  formatRound(c(2:8), 5) %>% 
  formatStyle(columns = c(1:8), 'text-align' = 'center')
```

#### Kd490 Pairwise comparison

```{r}
aa2 <-  kd_metrics %>% 
  filter(!is.na(Kd490)) %>% 
  filter(Metrics == 'mean') %>% 
  select(Kd490,Buffer_Size) %>% 
  mutate(Buffer_Size = Buffer_Size/1000)


pw2 <- pairwise.wilcox.test(aa2$Kd490,aa2$Buffer_Size)

pw2 %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('5',
                    '10',
                    '15',
                    '20'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  formatRound(c(2:5), 5) %>% 
  formatStyle(columns = c(1:5), 'text-align' = 'center')
```

### Figure 6 Kd490 and KdPAR, boxplot, 10km buffer, coastal hydrosystem types (10)

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(CLASS = case_when(
    CLASS == "Fjord" ~ 'Fiord',
    TRUE ~ CLASS)) %>% 
  mutate(CLASS = factor(CLASS, levels = c("Waituna-type lagoon", 
                                          "Hāpua-type lagoon", 
                                          "Beach stream",
                                          "Freshwater river mouth", 
                                          "Tidal river mouth", 
                                          "Tidal lagoon",
                                          "Shallow drowned valley",
                                          "Deep drowned valley",
                                          "Fiord"))) %>% 
  filter(Buffer_Size == 10000 & Metrics == 'mean') %>% 
  ggplot(data=.) +
  geom_boxplot(aes(x=factor(CLASS),y=KdPAR)) +
  scale_fill_viridis_d() +
  xlab('') +
  ylab('KdPAR (/m)') +
  theme_light() +
  theme(legend.position = 'bottom',
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


p2 <- kd_metrics %>% 
    mutate(CLASS = case_when(
    CLASS == "Fjord" ~ 'Fiord',
    TRUE ~ CLASS)) %>% 
    mutate(CLASS = factor(CLASS, levels = c("Waituna-type lagoon", 
                                          "Hāpua-type lagoon", 
                                          "Beach stream",
                                          "Freshwater river mouth", 
                                          "Tidal river mouth", 
                                          "Tidal lagoon",
                                          "Shallow drowned valley",
                                          "Deep drowned valley",
                                          "Fiord"))) %>% 
  filter(Buffer_Size == 10000 & Metrics == 'mean') %>% 
  ggplot(data=.) +
  geom_boxplot(aes(x=factor(CLASS),y=Kd490)) +
  scale_fill_viridis_d() +
  xlab('') +
  ylab('Kd490 (/m)') +
  theme_light() +
  theme(legend.position = 'bottom',
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

figure6 <- ggarrange(p2,
                     p,
                     labels = c("A", "B"),
                     ncol = 2, nrow = 1)
figure6
```

### KdPAR and Kd490 summary per class, 10km radius, and pairwise comparison

```{r}
#| message: false
#| warning: false

table2_kdPAR <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & Metrics == 'mean') %>% 
  filter(!is.na(KdPAR)) %>%
  group_by(CLASS) %>% 
  summarise(mean = mean(KdPAR,na.rm=T),
         sd = sd(KdPAR,na.rm=T),
         se = sd/sqrt(length(KdPAR))) 

DT::datatable(table2_kdPAR) %>% 
  formatRound(columns = c('mean','sd','se'), digits=3)
```

#### KdPAR Pairwise comparison

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>%   
  filter(Buffer_Size == 10000 & Metrics == 'mean') %>% 
  select(KdPAR,CLASS) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$CLASS)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('Waituna-type lagoon',
                    'Hāpua-type lagoon',
                    'Beach stream',
                    'Freshwater river mouth',
                    'Tidal river mouth',
                    'Tidal lagoon',
                    'Shallow drowned valley',
                    'Deep drowned valley'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  formatRound(c(2:9), 5) %>% 
  formatStyle(columns = c(1:9), 'text-align' = 'center')
```

```{r}
#| message: false
#| warning: false

table2_kd490 <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & Metrics == 'mean') %>% 
  filter(!is.na(Kd490)) %>%
  group_by(CLASS) %>% 
  summarise(mean = mean(Kd490,na.rm=T),
         sd = sd(Kd490,na.rm=T),
         se = sd/sqrt(length(Kd490))) 

DT::datatable(table2_kd490) %>% 
  formatRound(columns = c('mean','sd','se'), digits=3)
```

#### Kd490 Pairwise comparison

```{r}
#| message: false
#| warning: false

aa2 <- kd_metrics %>%   
  filter(Buffer_Size == 10000 & Metrics == 'mean') %>% 
  select(Kd490,CLASS) %>% na.omit()

pw2 <- pairwise.wilcox.test(aa2$Kd490,aa2$CLASS)

pw2 %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('Waituna-type lagoon',
                    'Hāpua-type lagoon',
                    'Beach stream',
                    'Freshwater river mouth',
                    'Tidal river mouth',
                    'Tidal lagoon',
                    'Shallow drowned valley'),
                  target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  formatRound(c(2:8), 5) %>% 
  formatStyle(columns = c(1:8), 'text-align' = 'center')
```

### Figure 7 Kd490 and KdPAR Boxplot, 10yr mean, 10km buffer, NEW grouping

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(CLASS2 = factor(CLASS2,levels = c('Open Coast','Enclosed Coast'))) %>% 
  filter(Buffer_Size == 10000 & Metrics == 'mean') %>% 
  ggplot(data=.) +
  geom_boxplot(aes(x=factor(CLASS2),y=KdPAR)) +
  scale_fill_viridis_d() +
  xlab('') +
  ylab('KdPAR (/m)') +
  theme_light() +
  theme(legend.position = 'bottom')

p2 <- kd_metrics %>% 
  mutate(CLASS2 = factor(CLASS2,levels = c('Open Coast','Enclosed Coast'))) %>% 
  filter(Buffer_Size == 10000 & Metrics == 'mean') %>% 
  ggplot(data=.) +
  geom_boxplot(aes(x=factor(CLASS2),y=Kd490)) +
  scale_fill_viridis_d() +
  xlab('') +
  ylab('Kd490 (/m)') +
  theme_light() +
  theme(legend.position = 'bottom')#axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)

figure7 <- ggarrange(p2,
                     p,
                     labels = c("A", "B"),
                     ncol = 2, nrow = 1)
figure7
```

#### KdPAR Pairwise comparison

```{r}
aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & Metrics == 'mean') %>% 
  select(KdPAR,CLASS2) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$CLASS2)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('Enclosed Coast'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  #formatRound(c(2), 5) %>% 
  formatStyle(columns = c(1:2), 'text-align' = 'center')
```

#### Kd490 Pairwise comparison

```{r}
aa2 <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & Metrics == 'mean') %>% 
  select(Kd490,CLASS2) 

pw2 <- pairwise.wilcox.test(aa2$Kd490,aa2$CLASS2)

pw2 %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('Enclosed Coast'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  formatRound(c(2), 3) %>% 
  formatStyle(columns = c(1:2), 'text-align' = 'center')
```

### Figure 8 Does stream order influence KdPAR? Analysis of 6 metrics

##### Main figure of case study

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & CLASS2 == "Open Coast" & Metrics %in% c('percentile10','median','percentile90')) %>% 
  mutate(Metrics = case_when(Metrics == 'percentile10' ~  '10th percentile',
                             Metrics == 'median' ~  'Median',
                             Metrics == 'percentile90' ~  '90th percentile')) %>%
    mutate(Metrics = factor(Metrics,levels = c('percentile5','10th percentile','Median','90th percentile','percentile95'))) %>% 
  ggplot(data=.,aes(x=factor(StreamOrde),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~Metrics,ncol=3, scales = 'free') +
  scale_fill_viridis_d() +
  ylab('KdPAR (/m)') + xlab('Stream Order') +
  ggtitle("Open Coast") +
  theme_light() +
  theme(legend.position = 'bottom')
p

p2 <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & CLASS2 == "Enclosed Coast" & Metrics %in% c('percentile10','median','percentile90')) %>% 
  mutate(Metrics = case_when(Metrics == 'percentile10' ~  '10th percentile',
                             Metrics == 'median' ~  'Median',
                             Metrics == 'percentile90' ~  '90th percentile')) %>%
    mutate(Metrics = factor(Metrics,levels = c('percentile5','10th percentile','Median','90th percentile','percentile95'))) %>% 
  ggplot(data=.,aes(x=factor(StreamOrde),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~Metrics,ncol=3, scales = 'free') +
  scale_fill_viridis_d() +
  ylab('KdPAR (/m)') + xlab('Stream Order') +
  ggtitle("Enclosed Coast") +
  theme_light() +
  theme(legend.position = 'bottom')
p2

figure8 <- ggarrange(p,
                     p2,
                     labels = c("A", "B"),
                     ncol = 1, nrow = 2)
figure8
```

##### Percentile 5

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & Metrics == 'percentile5') %>%
  ggplot(data=.,aes(x=factor(StreamOrde),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~CLASS2,scales='free',ncol=2) +
  scale_fill_viridis_d() +
  ylab('KdPAR 10yr (2013-2022) 5th percentile (/m)') + xlab('') +
  theme_light() +
  theme(legend.position = 'bottom')
p

```

###### Open Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
           Metrics == 'percentile5' &
         CLASS2 == 'Open Coast') %>% 
  select(KdPAR,StreamOrde) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$StreamOrde)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('5', '6','7'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  #formatRound(c(2), 5) %>% 
  formatStyle(columns = c(1:5), 'text-align' = 'center')
```

###### Enclosed Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
           Metrics == 'percentile5' &
         CLASS2 == 'Enclosed Coast') %>% 
  select(KdPAR,StreamOrde) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$StreamOrde)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('5', '6','7'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  #formatRound(c(2), 5) %>% 
  formatStyle(columns = c(1:5), 'text-align' = 'center')
```

##### Percentile 10

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & Metrics == 'percentile10') %>% 
  ggplot(data=.,aes(x=factor(StreamOrde),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~CLASS2,scales='free',ncol=2) +
  scale_fill_viridis_d() +
  ylab('KdPAR 10yr (2013-2022) 10th percentile (/m)') + xlab('') +
  theme_light() +
  theme(legend.position = 'bottom')
p

```

###### Open Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
           Metrics == 'percentile10' &
         CLASS2 == 'Open Coast') %>% 
  select(KdPAR,StreamOrde) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$StreamOrde)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('5', '6','7'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  #formatRound(c(2), 5) %>% 
  formatStyle(columns = c(1:5), 'text-align' = 'center')
```

###### Enclosed Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
           Metrics == 'percentile10' &
         CLASS2 == 'Enclosed Coast') %>% 
  select(KdPAR,StreamOrde) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$StreamOrde)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('5', '6','7'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  #formatRound(c(2), 5) %>% 
  formatStyle(columns = c(1:5), 'text-align' = 'center')
```

##### Mean

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & Metrics == 'mean') %>% 
  ggplot(data=.,aes(x=factor(StreamOrde),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~CLASS2,scales='free',ncol=2) +
  scale_fill_viridis_d() +
  ylab('KdPAR 10yr (2013-2022) Mean (/m)') + xlab('') +
  theme_light() +
  theme(legend.position = 'bottom')
p
```

###### Open Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
           Metrics == 'mean' &
         CLASS2 == 'Open Coast') %>% 
  select(KdPAR,StreamOrde) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$StreamOrde)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('5', '6','7'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  #formatRound(c(2), 5) %>% 
  formatStyle(columns = c(1:5), 'text-align' = 'center')
```

###### Enclosed Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
           Metrics == 'mean' &
         CLASS2 == 'Enclosed Coast') %>% 
  select(KdPAR,StreamOrde) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$StreamOrde)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('5', '6','7'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  #formatRound(c(2), 5) %>% 
  formatStyle(columns = c(1:5), 'text-align' = 'center')
```

##### Median

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & Metrics == 'median') %>% 
  ggplot(data=.,aes(x=factor(StreamOrde),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~CLASS2,scales='free',ncol=2) +
  scale_fill_viridis_d() +
  ylab('KdPAR 10yr (2013-2022) Median (/m)') + xlab('') +
  theme_light() +
  theme(legend.position = 'bottom')
p
```

###### Open Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
           Metrics == 'median' &
         CLASS2 == 'Open Coast') %>% 
  select(KdPAR,StreamOrde) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$StreamOrde)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('5', '6','7'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  #formatRound(c(2), 5) %>% 
  formatStyle(columns = c(1:5), 'text-align' = 'center')
```

###### Enclosed Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
           Metrics == 'median' &
         CLASS2 == 'Enclosed Coast') %>% 
  select(KdPAR,StreamOrde) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$StreamOrde)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('5', '6','7'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  #formatRound(c(2), 5) %>% 
  formatStyle(columns = c(1:5), 'text-align' = 'center')
```

##### Percentile 90

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & Metrics == 'percentile90') %>% 
  ggplot(data=.,aes(x=factor(StreamOrde),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~CLASS2,scales='free',ncol=2) +
  scale_fill_viridis_d() +
  ylab('KdPAR 10yr (2013-2022) 90th percentile (/m)') + xlab('') +
  theme_light() +
  theme(legend.position = 'bottom')
p
```

###### Open Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
           Metrics == 'percentile90' &
         CLASS2 == 'Open Coast') %>% 
  select(KdPAR,StreamOrde) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$StreamOrde)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('5', '6','7'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  #formatRound(c(2), 5) %>% 
  formatStyle(columns = c(1:5), 'text-align' = 'center')
```

###### Enclosed Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
           Metrics == 'percentile90' &
         CLASS2 == 'Enclosed Coast') %>% 
  select(KdPAR,StreamOrde) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$StreamOrde)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('5', '6','7'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  #formatRound(c(2), 5) %>% 
  formatStyle(columns = c(1:5), 'text-align' = 'center')
```

##### Percentile 95

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & Metrics == 'percentile95') %>% 
  ggplot(data=.,aes(x=factor(StreamOrde),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~CLASS2,scales='free',ncol=2) +
  scale_fill_viridis_d() +
  ylab('KdPAR 10yr (2013-2022) 95th percentile (/m)') + xlab('') +
  theme_light() +
  theme(legend.position = 'bottom')
p
```

###### Open Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
           Metrics == 'percentile95' &
         CLASS2 == 'Open Coast') %>% 
  select(KdPAR,StreamOrde) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$StreamOrde)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('5', '6','7'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  #formatRound(c(2), 5) %>% 
  formatStyle(columns = c(1:5), 'text-align' = 'center')
```

###### Enclosed Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
           Metrics == 'percentile95' &
         CLASS2 == 'Enclosed Coast') %>% 
  select(KdPAR,StreamOrde) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$StreamOrde)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(c('5', '6','7'),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  #formatRound(c(2), 5) %>% 
  formatStyle(columns = c(1:5), 'text-align' = 'center')
```

### Figure 9 Does Bioregions influence KdPAR? Analysis of 6 metrics

##### Main figure of case study

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & CLASS2 == "Open Coast" & Metrics %in% c('percentile10','median','percentile90')) %>% 
  mutate(Metrics = case_when(Metrics == 'percentile10' ~  '10th percentile',
                             Metrics == 'median' ~  'Median',
                             Metrics == 'percentile90' ~  '90th percentile')) %>%
    mutate(Metrics = factor(Metrics,levels = c('percentile5','10th percentile','Median','90th percentile','percentile95'))) %>% 
  ggplot(data=.,aes(x=factor(BioR__BioR),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~Metrics,ncol=3,scales = 'free') +
  scale_fill_viridis_d() +
  ylab('KdPAR (/m)') + xlab('Bioregions') +
  ggtitle("Open Coast") +
  theme_light() +
  theme(legend.position = 'bottom',
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p

p2 <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & CLASS2 == "Enclosed Coast" & Metrics %in% c('percentile10','median','percentile90')) %>% 
  mutate(Metrics = case_when(Metrics == 'percentile10' ~  '10th percentile',
                             Metrics == 'median' ~  'Median',
                             Metrics == 'percentile90' ~  '90th percentile')) %>%
    mutate(Metrics = factor(Metrics,levels = c('percentile5','10th percentile','Median','90th percentile','percentile95'))) %>% 
  ggplot(data=.,aes(x=factor(BioR__BioR),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~Metrics,ncol=3,scales = 'free') +
  scale_fill_viridis_d() +
  ylab('KdPAR (/m)') + xlab('Bioregions') +
  ggtitle("Enclosed Coast") +
  theme_light() +
  theme(legend.position = 'bottom',
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p2

figure9 <- ggarrange(p,
                     p2,
                     labels = c("A", "B"),
                     ncol = 1, nrow = 2)
figure9
```

##### Percentile 5

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & Metrics == 'percentile5') %>% 
  ggplot(data=.,aes(x=factor(BioR__BioR),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~CLASS2,scales='free',ncol=2) +
  scale_fill_viridis_d() +
  ylab('KdPAR 10yr (2013-2022) 5th percentile (/m)') + xlab('') +
  theme_light() +
  theme(legend.position = 'bottom',
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p
```

###### Open Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
         Metrics == 'percentile5' &
         CLASS2 == 'Open Coast') %>% 
  #mutate(CLASS2_BioR = paste0(CLASS2,"_",BioR__BioR)) %>% 
  select(KdPAR,BioR__BioR) 

#pw <- pairwise.wilcox.test(aa$KdPAR,aa$CLASS2_BioR)
pw <- pairwise.wilcox.test(aa$KdPAR,aa$BioR__BioR)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(columns = c(1:16),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  #formatRound(c(2), 5) %>% 
  formatStyle(columns = c(1:16), 'text-align' = 'center')
```

###### Enclosed Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
         Metrics == 'percentile5' &
         CLASS2 == 'Enclosed Coast') %>% 
  select(KdPAR,BioR__BioR) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$BioR__BioR)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(columns = c(1:16),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  #formatRound(c(2), 5) %>% 
  formatStyle(columns = c(1:16), 'text-align' = 'center')
```

##### Percentile 10

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & Metrics == 'percentile10') %>% 
  ggplot(data=.,aes(x=factor(BioR__BioR),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~CLASS2,scales='free',ncol=2) +
  scale_fill_viridis_d() +
  ylab('KdPAR 10yr (2013-2022) 10th percentile (/m)') + xlab('') +
  theme_light() +
  theme(legend.position = 'bottom',
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p
```

###### Open Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
         Metrics == 'percentile10' &
         CLASS2 == 'Open Coast') %>% 
  select(KdPAR,BioR__BioR) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$BioR__BioR)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(columns = c(1:16),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  formatStyle(columns = c(1:16), 'text-align' = 'center')
```

###### Enclosed Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
         Metrics == 'percentile10' &
         CLASS2 == 'Enclosed Coast') %>% 
  select(KdPAR,BioR__BioR) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$BioR__BioR)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(columns = c(1:16),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  formatStyle(columns = c(1:16), 'text-align' = 'center')
```

##### Mean

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & Metrics == 'mean') %>% 
  ggplot(data=.,aes(x=factor(BioR__BioR),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~CLASS2,scales='free',ncol=2) +
  scale_fill_viridis_d() +
  ylab('KdPAR 10yr (2013-2022) Mean (/m)') + xlab('') +
  theme_light() +
  theme(legend.position = 'bottom',
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p
```

###### Open Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
         Metrics == 'mean' &
         CLASS2 == 'Open Coast') %>% 
  select(KdPAR,BioR__BioR) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$BioR__BioR)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(columns = c(1:16),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  formatStyle(columns = c(1:16), 'text-align' = 'center')
```

###### Enclosed Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
         Metrics == 'mean' &
         CLASS2 == 'Enclosed Coast') %>% 
  select(KdPAR,BioR__BioR) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$BioR__BioR)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(columns = c(1:16),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  formatStyle(columns = c(1:16), 'text-align' = 'center')
```

##### Median

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & Metrics == 'median') %>% 
  ggplot(data=.,aes(x=factor(BioR__BioR),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~CLASS2,scales='free',ncol=2) +
  scale_fill_viridis_d() +
  ylab('KdPAR 10yr (2013-2022) Median (/m)') + xlab('') +
  theme_light() +
  theme(legend.position = 'bottom',
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p
```

###### Open Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
         Metrics == 'median' &
         CLASS2 == 'Open Coast') %>% 
  select(KdPAR,BioR__BioR) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$BioR__BioR)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(columns = c(1:16),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  formatStyle(columns = c(1:16), 'text-align' = 'center')
```

###### Enclosed Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
         Metrics == 'median' &
         CLASS2 == 'Enclosed Coast') %>% 
  select(KdPAR,BioR__BioR) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$BioR__BioR)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(columns = c(1:16),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  formatStyle(columns = c(1:16), 'text-align' = 'center')
```

##### Percentile 90

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & Metrics == 'percentile90') %>% 
  ggplot(data=.,aes(x=factor(BioR__BioR),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~CLASS2,scales='free',ncol=2) +
  scale_fill_viridis_d() +
  ylab('KdPAR 10yr (2013-2022) 90th percentile (/m)') + xlab('') +
  theme_light() +
  theme(legend.position = 'bottom',
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p
```

###### Open Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
         Metrics == 'percentile90' &
         CLASS2 == 'Open Coast') %>% 
  select(KdPAR,BioR__BioR) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$BioR__BioR)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(columns = c(1:16),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  formatStyle(columns = c(1:16), 'text-align' = 'center')
```

###### Enclosed Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
         Metrics == 'percentile90' &
         CLASS2 == 'Enclosed Coast') %>% 
  select(KdPAR,BioR__BioR) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$BioR__BioR)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(columns = c(1:16),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  formatStyle(columns = c(1:16), 'text-align' = 'center')
```

##### Percentile 95

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & Metrics == 'percentile95') %>% 
  ggplot(data=.,aes(x=factor(BioR__BioR),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~CLASS2,scales='free',ncol=2) +
  scale_fill_viridis_d() +
  ylab('KdPAR 10yr (2013-2022) 95th percentile (/m)') + xlab('') +
  theme_light() +
  theme(legend.position = 'bottom',
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p
```

###### Open Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
         Metrics == 'percentile95' &
         CLASS2 == 'Open Coast') %>% 
  select(KdPAR,BioR__BioR) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$BioR__BioR)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(columns = c(1:16),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  formatStyle(columns = c(1:16), 'text-align' = 'center')
```

###### Enclosed Coast

```{r}
#| message: false
#| warning: false

aa <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & 
         Metrics == 'percentile95' &
         CLASS2 == 'Enclosed Coast') %>% 
  select(KdPAR,BioR__BioR) 

pw <- pairwise.wilcox.test(aa$KdPAR,aa$BioR__BioR)

pw %>% 
  tidy() %>% 
  pivot_wider(names_from = c(group2),values_from = p.value) %>% 
  DT::datatable() %>% 
  DT::formatStyle(columns = c(1:16),target = 'cell',fontWeight = styleInterval(.05, c('bold', 'normal'))) %>% 
  formatStyle(columns = c(1:16), 'text-align' = 'center')
```

### Figure 10 Time series of National and river with highest KdPAR

#### Time series national scale

```{r}
#| message: false
#| warning: false

kd_ts_10km <- read_csv('KdPAR_SCENZ_1D_v05_Buffer_10000m_RiverMouths_REC2.csv')

kd_ts_10km <- inner_join(kd_ts_10km,rivermouth %>% select(RID,NAME_name2,CLASS2))

kd_ts_10km_yearly_sum <- kd_ts_10km %>% 
  mutate(year = year(Date)) %>% 
  group_by(year,CLASS2) %>% 
  summarise(KdPAR_median = median(KdPAR,na.rm=T),
            KdPAR_10p = quantile(KdPAR,na.rm=T,0.1),
            KdPAR_90p = quantile(KdPAR,na.rm=T,0.9))
  
p <- kd_ts_10km_yearly_sum %>% 
  filter(CLASS2 == 'Open Coast') %>% 
  pivot_longer(-c(year,CLASS2)) %>% 
  mutate(name = case_when(name == 'KdPAR_10p' ~  '10th percentile',
                             name == 'KdPAR_median' ~  'Median',
                             name == 'KdPAR_90p' ~  '90th percentile')) %>%
  mutate(name = factor(name,levels = c('10th percentile','Median','90th percentile'))) %>% 
  ggplot(data=.,aes(x=year,y=value)) +
  geom_point() +
  geom_line() + 
  ylab('KdPAR (/m)') + xlab('Year') +
  geom_smooth(method = 'lm') +
  ggtitle("Open Coast") +
  facet_wrap(~name,nrow=1,scales = 'free') +
  theme_light()
p

p2 <- kd_ts_10km_yearly_sum %>% 
  filter(CLASS2 == 'Enclosed Coast') %>% 
  pivot_longer(-c(year,CLASS2)) %>% 
  mutate(name = case_when(name == 'KdPAR_10p' ~  '10th percentile',
                             name == 'KdPAR_median' ~  'Median',
                             name == 'KdPAR_90p' ~  '90th percentile')) %>%
  mutate(name = factor(name,levels = c('10th percentile','Median','90th percentile'))) %>% 
  ggplot(data=.,aes(x=year,y=value)) +
  geom_point() +
  geom_line() + 
  ylab('KdPAR (/m)') + xlab('Year') +
  geom_smooth(method = 'lm') +
  ggtitle("Enclosed Coast") +
  facet_wrap(~name,nrow=1,scales = 'free') +
  theme_light()
p2
figure10 <- ggarrange(p,
                     p2,
                     labels = c("A", "B"),
                     ncol = 1, nrow = 2)
figure10
```

##### Trends

```{r}
trends <- kd_ts_10km_yearly_sum %>% 
  pivot_longer(-c(year,CLASS2)) %>% 
  group_by(CLASS2,name) %>% 
  nest() %>% 
    mutate(ts_out = purrr::map(data, ~ts(.x$value,start=2002,end=2023,frequency = 1))) %>% 
  mutate(sens = purrr::map(ts_out, ~sens.slope(.x, conf.level = 0.95))) %>% 
  mutate(Sens_Slope = as.numeric(unlist(sens)[1]),P_Value =as.numeric(unlist(sens)[3])) %>% 
  select(CLASS2,name, Sens_Slope, P_Value)
  
trends %>% 
  DT::datatable() 
```

#### Rivermouths ranked by highest KdPAR values (mean 2002-2023 period)

```{r}
## Rankings of most dirty rivermouth
kd_ts_10km_summary <- kd_ts_10km %>% 
  group_by(RID,NAME,name_2,CLASS2) %>% 
  summarise(KdPAR_mean = round(mean(KdPAR,na.rm=T),3),
            KdPAR_sd = round(sd(KdPAR,na.rm=T),3)) %>% 
  arrange(desc(KdPAR_mean))

#write_csv(kd_ts_10km_summary,'KdPAR_2002_2023_mean_sd_Rivermouth_ranked.csv')

kd_ts_10km_summary %>% 
  DT::datatable() 
```

### Figure 11 KdPAR mean for all rivers

```{r}
kd_mean_10km <- kd_ts_10km %>% group_by(CLASS2,RID,NAME,name_2,NAME_name2) %>% 
  summarise(KdPAR_mean = mean(KdPAR,na.rm=T)) %>% 
  arrange(desc(KdPAR_mean))

p <- kd_mean_10km %>%
  arrange(desc(KdPAR_mean)) %>% 
  ggplot(data=.,aes(x=fct_inorder(NAME_name2) ,y=KdPAR_mean,fill=CLASS2)) +#colour
  #geom_segment(aes(xend=NAME_name2 ,y = 0, yend=KdPAR_mean,colour=CLASS2),size=.5) +
  #scale_colour_manual(values=c('blue','red')) +
  #geom_point(size=2) +
  scale_fill_manual(values=c('blue','red')) +
  geom_col() +
  xlab('Rivers (n = 226)') + ylab('KdPAR mean (/m)') +
  theme_classic() +
  theme(legend.position = 'bottom',
        legend.title = element_blank(),
        axis.text.x=element_blank(), 
        axis.ticks.x=element_blank())
p

p2 <- kd_mean_10km[1:20,] %>%# kd_mean_10km[1:20,]
  ggplot(data=.,aes(x=fct_inorder(name_2) ,y=KdPAR_mean,fill=CLASS2)) +#colour
  #geom_segment(aes(xend= ,y = 0, yend=KdPAR_mean,colour=CLASS2)) +
  #scale_colour_manual(values=c('blue','red')) +
  scale_fill_manual(values=c('blue','red')) +
  geom_col() +
  #geom_point(size=2) +
  xlab('Rivers (top 20)') + ylab('KdPAR mean (/m)') +
  theme_light() +
  theme(legend.position = 'none',
        axis.text.x = element_text(size = 12, angle = 90, vjust = 0.5, hjust=1))
p2

figure11 <- ggarrange(p,
                     p2,
                     widths = c(5, 5),#3,2
                     labels = c("A", "B"),
                     ncol = 2, nrow = 1)
figure11
```

### Supplementary Information

### Fig S1

```{r}
#| message: false
#| warning: false


figS1 <- kd_metrics %>% 
  filter(Metrics %in% c('median','percentile10','percentile90')) %>% 
  filter(!is.na(KdPAR)) %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  mutate(Metrics = case_when(Metrics == 'percentile10' ~  '10th percentile',
                             Metrics == 'median' ~  'Median',
                             Metrics == 'percentile90' ~  '90th percentile')) %>%
  mutate(Metrics = factor(Metrics,levels = c('10th percentile','Median','90th percentile'))) %>% 
  ggplot(data=.,aes(x=factor(Buffer_Size),y=KdPAR,fill=Metrics)) +
  geom_boxplot() +
  scale_fill_viridis_d() +
  facet_wrap(~CLASS) +
  #ggtitle('Percentage of valid observations of KdPAR (2013-2022)') +
  ylab('KdPAR (/m)') + xlab('Radius (km)') +
  theme_light() +
  theme(legend.position = 'bottom')
figS1
```

### Fig S2

```{r}
#| message: false
#| warning: false

figS2 <- kd_metrics %>% 
  filter(Metrics %in% c('median','percentile10','percentile90')) %>% 
  filter(!is.na(KdPAR)) %>% 
  mutate(CLASS2 = factor(CLASS2,levels = c('Open Coast','Enclosed Coast'))) %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  mutate(Metrics = case_when(Metrics == 'percentile10' ~  '10th percentile',
                             Metrics == 'median' ~  'Median',
                             Metrics == 'percentile90' ~  '90th percentile')) %>%
  mutate(Metrics = factor(Metrics,levels = c('10th percentile','Median','90th percentile'))) %>% 
  ggplot(data=.,aes(x=factor(Buffer_Size),y=KdPAR,fill=Metrics)) +
  geom_boxplot() +
  scale_fill_viridis_d() +
  facet_wrap(~CLASS2) +
  #ggtitle('Percentage of valid observations of KdPAR (2013-2022)') +
  ylab('KdPAR (/m)') + xlab('Radius (km)') +
  theme_light() +
  theme(legend.position = 'bottom')
figS2
```

### Fig S3 KdPAR 90th percentile vs Catchment Size

```{r}
p <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & Metrics == 'percentile90' & CATCHMT_AR>0 & CLASS2 == 'Open Coast') %>% 
  ggplot(data=.,aes(x=CATCHMT_AR,y=KdPAR)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ylab('KdPAR 10yr (2013-2022) 90th percentile (/m)') + xlab('Catchment Size (km2)') +
  ggtitle('Open Coast') +
  theme_light() +
  theme(legend.position = 'bottom')
p

p2 <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & Metrics == 'percentile90' & CATCHMT_AR>0 & CLASS2 == 'Enclosed Coast') %>% 
  ggplot(data=.,aes(x=CATCHMT_AR,y=KdPAR)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ylab('KdPAR 10yr (2013-2022) 90th percentile (/m)') + xlab('Catchment Size (km2)') +
  ggtitle('Enclosed Coast') +
  theme_light() +
  theme(legend.position = 'bottom')
p2

figS3 <- ggarrange(p,
                     p2,
                     labels = c("A", "B"),
                     ncol = 2, nrow = 1)
figS3
```

### Fig S4 Same as Figure 8 but for coastal council regions

```{r}
#| message: false
#| warning: false

p <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & CLASS2 == "Open Coast" & Metrics %in% c('percentile10','median','percentile90')) %>% 
  mutate(Metrics = case_when(Metrics == 'percentile10' ~  '10th percentile',
                             Metrics == 'median' ~  'Median',
                             Metrics == 'percentile90' ~  '90th percentile')) %>%
    mutate(Metrics = factor(Metrics,levels = c('percentile5','10th percentile','Median','90th percentile','percentile95'))) %>% 
  ggplot(data=.,aes(x=factor(REGC2017_N),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~Metrics,ncol=3,scales = 'free') +
  scale_fill_viridis_d() +
  ylab('KdPAR (/m)') + xlab('') +
  ggtitle("Open Coast") +
  theme_light() +
  theme(legend.position = 'bottom',
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p

p2 <- kd_metrics %>% 
  mutate(Buffer_Size = Buffer_Size/1000) %>% 
  filter(Buffer_Size == 10 & !is.na(KdPAR) & CLASS2 == "Enclosed Coast" & Metrics %in% c('percentile10','median','percentile90')) %>% 
  mutate(Metrics = case_when(Metrics == 'percentile10' ~  '10th percentile',
                             Metrics == 'median' ~  'Median',
                             Metrics == 'percentile90' ~  '90th percentile')) %>%
    mutate(Metrics = factor(Metrics,levels = c('percentile5','10th percentile','Median','90th percentile','percentile95'))) %>% 
  ggplot(data=.,aes(x=factor(REGC2017_N),y=KdPAR)) +
  geom_boxplot() +
  facet_wrap(~Metrics,ncol=3,scales = 'free') +
  scale_fill_viridis_d() +
  ylab('KdPAR (/m)') + xlab('') +
  ggtitle("Enclosed Coast") +
  theme_light() +
  theme(legend.position = 'bottom',
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p2

figureS4 <- ggarrange(p,
                     p2,
                     labels = c("A", "B"),
                     ncol = 1, nrow = 2)
figureS4
```

### Figure S5 Number of valid observations per class2, 10 km radius

```{r}
#| message: false
#| warning: false
p <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & Metrics == 'valid_days_pc') %>% 
  mutate(CLASS2 = factor(CLASS2,levels = c('Open Coast','Enclosed Coast'))) %>% 
  ggplot(data=.) +
  geom_boxplot(aes(x=factor(CLASS2),y=KdPAR)) +
  scale_fill_viridis_d() +
  ggtitle("500 m KdPAR product") +
  ylab('Percent of valid observations (%)') + xlab('') +
  theme_light() +
  theme(legend.position = 'bottom',
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

p2 <- kd_metrics %>% 
  filter(Buffer_Size == 10000 & Metrics == 'valid_days_pc') %>% 
  mutate(CLASS2 = factor(CLASS2,levels = c('Open Coast','Enclosed Coast'))) %>% 
  ggplot(data=.) +
  geom_boxplot(aes(x=factor(CLASS2),y=Kd490)) +
  scale_fill_viridis_d() +
  ggtitle("4 km Kd490 product") +
  ylab('Percent of valid observations (%)') + xlab('') +
  theme_light() +
  theme(legend.position = 'bottom',
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

figureS5 <- ggarrange(p2,
                     p,
                     labels = c("A", "B"),
                     ncol = 2, nrow = 1)
figureS5
```
